---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Tabular Data {#tabular-data}

- Aliases: 
  + Tabular files
  + Flat
  + Delimited
- Includes:
  + Comma-Separated Value (.csv)
  + Tab-Separated Value (.tsv)

## Example Data

Modified from http://www.gapminder.org/data/

```
country,continent,year,lifeExp,pop,gdpPercap       # header/column names
Afghanistan,Asia,1952,28.801,8425333,779.4453145
Afghanistan,Asia,1957,30.332,9240934,820.8530296
Afghanistan,Asia,1962,31.997,10267083,853.10071
Afghanistan,Asia,1967,34.02,11537966,836.1971382
Afghanistan,Asia,1972,36.088,13079460,739.9811058
Afghanistan,Asia,1977,38.438,14880372,786.11336
Afghanistan,Asia,1982,39.854,12881816,978.0114388
Afghanistan,Asia,1987,40.822,13867957,852.3959448
Afghanistan,,,N/A,,                                # notice that we're missing values
```

```{r}
csv_text <- 
'country,continent,year,lifeExp,pop,gdpPercap
Afghanistan,Asia,1952,28.801,8425333,779.4453145
Afghanistan,Asia,1957,30.332,9240934,820.8530296
Afghanistan,Asia,1962,31.997,10267083,853.10071
Afghanistan,Asia,1967,34.02,11537966,836.1971382
Afghanistan,Asia,1972,36.088,13079460,739.9811058
Afghanistan,Asia,1977,38.438,14880372,786.11336
Afghanistan,Asia,1982,39.854,12881816,978.0114388
Afghanistan,Asia,1987,40.822,13867957,852.3959448
Afghanistan,,,N/A,,'

csv_file <- tempfile(fileext = ".csv")
writeLines(text = csv_text, con = csv_file)
```


## Input: Reading Data

```{r}
library(readr)
```


### Default

```{r}
data_frame_from_csv <- read_csv(file = csv_file)
data_frame_from_csv
```

#### Diagnose Problems

Notice that our `year` column says `<dbl>`, referring it being of type `double`, yet all of our `year` values are whole numbers.

```{r}
typeof(data_frame_from_csv$year)
data_frame_from_csv$year
```

We also hanve `"N/A"` in our `lifeExp` column, forcing R to interpret all `lifeExp` values as `character`s (`<chr>`).

```{r}
typeof(data_frame_from_csv$lifeExp)
data_frame_from_csv$lifeExp
```

#### Solution

```{r}
read_csv(
  file = csv_file,
  col_types = cols(
    country = col_character(),
    continent = col_character(),
    year = col_integer(),        # read `year` as `integer`
    lifeExp = col_double(),      # read `lifeExp` as `double`
    pop = col_double(),
    gdpPercap = col_double()
  ),
  na = c("", "N/A")              # be explicit about how missing values may be represented
)
```


