---
output: html_document
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE}
options(width=300)
```


# Manipulating Data Frames {#manipulating-data-frames}


```{r}
library(tidyverse, warn.conflicts = FALSE)

df <- tibble(
  group = c("a", "a", "b", "b", "b"),
  a = c(1, 4, NA, 3, 5),
  b = c(9, NA, 8, 10, 7),
  c = c(TRUE, FALSE, NA, FALSE, TRUE),
  d = c(LETTERS[1:3], NA, LETTERS[[5]]),
  e = factor(1:5, labels = c("tiny", "small", "medium", "big", "huge")),
  f_col = c(as.Date(NA), as.Date("2020-09-23") + c(3, 2, 1, 4)),
  g_col = c(as.POSIXct("2020-09-23 00:00:00") + 1:4 * 60 * 60 * 24 * 1.1, NA),
  col_h = list(c(1, 10), c(2, NA), c(3, 8), c(4, 7), c(5, 6)),
  col_i = list(NULL, pi, month.abb[6:10], iris, as.matrix(mtcars))
)

df
glimpse(df)
```

## `select()` Columns



### by Name

```{r}
df %>%
  select(a)
```

```{r}
df %>%
  select(a, c, e)
```

```{r}
df %>%
  select(b, d, f_col)
```

```{r}
df %>%
  select(b, c, everything())
```

```{r}
df %>%
  select(b, c, everything(), -a)
```

```{r}
cols_to_select <- c("a", "c", "e")
df %>%
  select(all_of(cols_to_select))
```


<!-- ```{r, error=TRUE} -->
<!-- cols_to_select <- c("a", "c", "e", "missing_column") -->
<!-- df %>% -->
<!--   select(all_of(cols_to_select)) -->
<!-- ``` -->

<!-- ```{r, error=TRUE} -->
<!-- cols_to_select <- c("a", "c", "e", "missing_column") -->
<!-- df %>% -->
<!--   select(any_of(cols_to_select)) -->
<!-- ``` -->



### by Index

```{r}
df %>%
  select(1L)
```

```{r}
df %>%
  select(1, 3, 5)
```


```{r}
df %>%
  select(2, 4, 6)
```


```{r}
df %>%
  select(2:3, everything())
```

```{r}
df %>%
  select(2:3, everything(), -1)
```


```{r}
cols_to_select <- c(1, 3, 5)
df %>%
  select(all_of(cols_to_select))
```


<!-- ```{r, error=TRUE} -->
<!-- cols_to_select <- c(1, 3, 5, 1000) -->
<!-- df %>% -->
<!--   select(all_of(cols_to_select)) -->
<!-- ``` -->


```{r}
cols_to_select <- c(1, 3, 5, 1000)
df %>%
  select(any_of(cols_to_select))
```


### by Name Pattern

`contains()` selects a column if _any_ part of its name contains `match=`.

```{r}
df %>%
  select(contains(match = "col"))
```

`starts_with()` selects a column if its name starts with `match=`.

```{r}
df %>%
  select(starts_with("col_"))
```

`starts_with()` selects a column if its name ends with `match=`.

```{r}
df %>%
  select(ends_with("_col"))
```

`matches()`s Selects a column if its name matches a regular expression pattern.

```{r}
df %>%
  select(matches("(^\\w_)?col(_\\w)?"))
```


### by Data Type

```{r}
df %>%
  select(where(is.factor))
```

```{r}
df %>%
  select_if(is.factor)
```

```{r}
df %>%
  select(where(is.factor), f_col)
```


```{r}
df %>%
  select(a, !where(is.integer))
```



```{r}
df %>%
  select(where(is.character) | where(is.factor))
```

```{r}
df %>%
  select(where(~ is.double(.) | is.list(.)))
```

```{r}
df %>%
  select_if(~ is.character(.x) | is.factor(.x))
```

## `filter()` Rows

### by `row_number()`

```{r}
df %>%
  filter(row_number() == 1)
```

```{r}
df %>%
  filter(row_number() > 1)
```


### by Name

```{r}
df %>%
  filter(a == 2)
```

```{r}
df %>%
  filter(a != 2)
```

```{r}
df %>%
  filter(c)
```

```{r}
df %>%
  filter(!c)
```


```{r}
df %>%
  filter(a == 5, d == "E")
```

```{r}
df %>%
  filter(a >= 3 | f_col == "2020-09-24")
```

```{r}
df %>%
  filter(a < 2 | c)
```

```{r}
df %>%
  filter(!is.na(a), !is.na(b), !is.na(d))
```

### by Type

```{r}
df %>%
  filter(across(where(is.numeric), ~ .x >= 5))
```


```{r}
df %>%
  filter_if(is.numeric, ~ .x >= 5)
```


```{r}
df %>%
  filter_if(is.list, ~ map_lgl(.x, ~ !is.null(.x)))
```


## `arrange()` Rows

```{r}
df %>%
  arrange(a)
```

```{r}
df %>%
  arrange(desc(a))
```



